### 2018-04-11 paper.Rmd
streamlined teaching section

```{r load_packages, echo = FALSE, warning=FALSE}
library(papaja)
library(tidyverse)
library(childesr)
library(directlabels)
library(koRpus)
library(here)
knitr::opts_chunk$set(cache = TRUE, echo = FALSE)
```

@@

<!-- [[Details of extraction]] -->
<!-- [[mention using XML childes representation]] -->
<!-- [[Details of testing: Why trust our parse over your in-house library? CLAN TESTING DETAILS - also mention versioning as a reason to trust]] -->

@@

- Sparsely annotated tiers, e.g. phonology (`%pho`) and situation (`%sit`)
- Media links, including timing information
- compound word markers, indicated with underscores ("green_eggs_and_ham")
- Tone direction and stress
- Filled pauses 
- Reformulations, word revision, and phrase revision, e.g. <what did you> [//] how can you see it ?
- paralinguistic material, e.g. [=! cries]

@@

Teachers of courses on early language acquisition often want to demonstrate interesting changes in children’s early language. One teaching method is to present static displays on slides or handouts that show (a) text from parent-child conversations extracted from CHILDES or (b) data visualizations of changes in various metrics related to language production and input (e.g., MLU or Frequency). The challenge with static displays or handouts is that they cannot be modified during a lecture and thus rely on the instructor selecting examples that they think will be compelling to their students.

Interactive, in-class demonstrations offer a solution and can be a powerful way to explain complex concepts while increasing student engagement with the course materials. One goal in creating the browser-based interactive visualization tools in childes-db is to provide instructors with a new tool for illustrating key concepts in children's language acquisition. In this section, we walk through one example using childes-db to show the developmental trajectories of different kinds of words that in children's early production. We also highlight ways that instructors could make the in-class demonstration interactive by eliciting hypotheses from students, which become operationalized as queries to the interactive visualization tools.

### Worked example

What kinds of words do children first produce? This question is a good candidate for an interesting "hook" that gets students thinking about some key issues in child language acquisition. Diary studies and large-scale studies using parent report show that children’s first words tend to fall into a fairly small number of categories: people, food, body parts, clothing, animals, vehicles, toys, household objects, routines, and activities or states [@clark2009first; @fenson1994variability]. The key insight is that young children talk about what is going on around them: people they see every day, e.g., toys and small household objects they can manipulate or food they can control.

A teacher could illustrate children's early vocabularies by showing a table with columns for the lexical category, examples of words from that category, and children's average age of production. To make this an interactive demonstration using childes-db, the teacher could:

1. introduce the question: What kinds of words do children first produce?
2. allow students to reflect or do a pair-and-share discussion with their neighbor
3. show the trajectory of a single lexical item while explaining the key parts of the visualization (see Panel A of Fig. XX)
4. elicit hypotheses from students about the kinds of words that children are likely to produce 
5. make real-time queries to the interactive visualization tool to add students' suggestions and talk through the updated plots 
6. finish by entering a pre-selected set of words that communicate the critical takeaway (e.g., see Panel C of Fig. XX)

```{r in-class-demo-flow, fig.pos="h", out.width = "100%", fig.cap="Worked example of using the interactive visualization tools for in-class teaching. Panels A-D show how an instructor could dynamically build a plot during a lecture to interactively demonstrate a key concept in language acquisition."}
library(here)
library(png)
png::readPNG(here::here("imgs/in_class_teach_viz.png")) %>% grid::grid.raster()
```

Another benefit of this workflow is that the instructor could easily reuse it to demonstrate other concepts or questions that come up during the lecture, e.g., order of production for different types of words such as quantifiers vs. number words or function vs. content words. Moreover, future work on childes-db interactive visualization tools could add part of speech tags, phonological information, and n-grams. This would allow instructors to use childes-db to demonstrate concepts such as when children start producing certain phonological patterns (e.g., consonant clusters), the debate about the extent of the noun/verb bias in children's early vocabulary, and the development of particular syntactic constructions.

In addition to the interactive data visualizations, instructors can use childes-db to download the raw data used to generate the figures. This allows teachers an opportunity to have students further explore and analyze the data they see in the plots. Moreover, this functionality provides an example of another way childes-db can facilitate language acquisition courses by making data easily-accessible for tutorials and student research projects. In the next section, we walk through one example of how an instructor could use childes-db to create a tutorial that scaffolds students' ability to ask interesting computational questions using child language data.

@@

In the introduction, we outlined the challenge of getting students started with analyzing data from CHILDES. These barriers include learning the details of a special file format (CHAT) and a command-line tool (CLAN) built to specifically for the CHAT format. This workflow can be daunting to navigate for a beginner. Thus, an instructor would have to spend course time teaching students to use CHAT and CLAN, leaving less time for students to try analyses that might increase their interest in child language acquisition. Moreover,  while CLAN provides a useful tool for interactive exploration,  it does not connect to a pipeline of modern natural language processing tools [e.g., models of word embeddings; @mikolov2013distributed], making it difficult for students to explore the data using new NLP tools.

We hope that childes-db combined with the `childesr` API will provide instructors with a solution to both of these challenges. The API gives students an easy-to-learn method for programmatic access to CHILDES data. This will reduce the time spent getting data out of CHILDES and allow more time for thinking about interesting questions that can be answered with NLP and child language data. Second, by writing our API in the R programming language, students will gain access to a general workflow can be integrated into a pipeline that includes the newest NLP tools.

For example, an instructor could create a programming assignment with the goal of reproducing the key findings in the case studies that we presented earlier -- color words or gender. Depending on the students' knowledge of R, the instructor could decide how much of the `childesr` code to provide before asking students to generate their own plots and writeups.

Moreover, the instructor could take advantage of recent advances in tools for teaching literate programming such as R Markdown [@xie2015dynamic; @baumer2014r]. Writing the assignment in an R Markdown notebook would allow the interleaving of prose with executable chunks of code. Instructors could then write prose that is directly related to the course content and the assignment while building in pre-written blocks of executable code that demonstrate key programming concepts for students. To push students' understanding, instructors could write incomplete or blank code blocks that students can use to write their own code. In short, `childesr` combined with R Markdown provides a powerful teaching tool, allowing students to start analyzing language data quickly and providing instructors with a way to build customized lessons for students at different levels of programming experience.

>>>>

```{r load_packages, echo = FALSE, warning=FALSE}
library(papaja)
library(tidyverse)
library(childesr)
library(directlabels)
library(here)
knitr::opts_chunk$set(cache = TRUE, echo = FALSE)
```

@@

In order to import the CHILDES corpora into the MySQL schema described above, it must first be accurately parsed and subsequently vetted to ensure the integrity of the data migration. Along with being made available in CHAT, the corpora are also downloadable in a format known as eXtensible Markup Language, or XML; this significantly mitigated any problems related to parsing CHAT as it allowed us to leverage the built-in XML parsing libraries available in numerous programming languages such as Python. Once the XML data were crawled and parsed into tokens and utterances, they were imported into a MySQL database. The meticulously constructed CHILDES XML Schema allowed us to make precise decisions as to what information we found relevant to our database and what was to be discarded or ignored; further information on what data were retained can be found below.

The data imported into childes-db requires extensive data integrity checks to ensure that our parse of the corpora is accurate and preferable over the adhoc parsers developed by many individual researchers. In order to evaluate this, we compared unigram counts in our database with those outputted by CLAN, the command-line tool built specifically for analysis of transcripts coded in CHAT. <!-- technically this is all we do right now -->

<!-- [[Details of testing: Why trust our parse over your in-house library? CLAN TESTING DETAILS - also mention versioning as a reason to trust]] -->

@@

- Sparsely annotated tiers, e.g. phonology (`%pho`) and situation (`%sit`)
- Media links, including timing information <!-- we have media timestamps -->
- compound word markers, indicated with underscores ("green_eggs_and_ham") <!-- we have this -->
- Tone direction and stress
- Filled pauses 
- Reformulations, word revision, and phrase revision, e.g. <what did you> [//] how can you see it ?
- paralinguistic material, e.g. [=! cries]

@@

Teachers of courses on early language acquisition often want to demonstrate interesting changes in children’s early language. One method is to present static displays that show text from parent-child conversations extracted from CHILDES or data visualizations of various metrics of production and input (e.g., MLU or Frequency). The challenge with static displays is that they cannot be modified during a lecture and thus rely on the instructor selecting examples that will be compelling to students.

In-class demonstrations offer a solution and can be a powerful way to explain complex concepts while increasing student engagement with the course materials. One goal of creating the web applications in childes-db is to give instructors a new tool for illustrating concepts in children's language acquisition.

For example, a teacher could want to show the types of words that children first produce. Diary studies and large-scale studies using parent report show that children’s first words tend to fall into a fairly small number of categories: people, food, body parts, clothing, animals, vehicles, toys, household objects, routines, and activities or states [@clark2009first; @fenson1994variability]. The key insight is that young children talk about what is going on around them: people they see every day, e.g., toys and small household objects they can manipulate or food they can control.

A teacher could illustrate children's early vocabularies by showing a table with columns for the lexical category, examples of words from that category, and children's average age of production. To make this interactive using the web apps, the teacher could:

1. introduce the research question
2. allow students to reflect or do a pair-and-share discussion with their neighbor
3. show the trajectory of a single lexical item while explaining the key parts of the visualization (see Panel A of Figure \ref{fig:in-class-demo-plot})
4. elicit hypotheses from students about the kinds of words that children are likely to produce 
5. make real-time queries to the web applications to add students' suggestions and talk through the updated plots (Panels B and C of Figure \ref{fig:in-class-demo-plot})
6. finish by entering a pre-selected set of words that communicate the critical takeaway (Panel D of Figure \ref{fig:in-class-demo-plot})

```{r in-class-demo-plot, fig.pos="h", out.width = "100%", fig.cap="Worked example of using the web applications for in-class teaching. Panels A-D show how an instructor could dynamically build a plot during a lecture to interactively demonstrate a key concept in language acquisition."}
library(here)
library(png)
png::readPNG(here::here("imgs/in_class_teach_viz.png")) %>% grid::grid.raster()
```

In addition to the web applications, instructors can use childes-db to download the raw data used to generate the figures. This allows teachers an opportunity to have students further explore and analyze the data they see in the plots. Moreover, this functionality provides an example of another way childes-db can facilitate language acquisition courses by making data easily-accessible for tutorials and student research projects.

@@

One goal of childes-db and `childesr` API is to provide instructors with an easy-to-learn method for giving students programmatic access to CHILDES data. This will reduce the time spent getting data out of CHILDES and allow more time for thinking about interesting questions that can be answered with NLP and child language data. Second, by writing our API in the R programming language, students will gain access to a general workflow can be integrated into a pipeline that includes the newest NLP tools.

For example, an instructor could create a programming assignment with the goal of reproducing the key findings in the case studies presented above -- color words or gender. Depending on the students' knowledge of R, the instructor could decide how much of the `childesr` starter code to provide before asking students to generate their own plots and writeups. Moreover, the instructor could take advantage of recent advances in tools for teaching literate programming such as R Markdown [@xie2015dynamic; @baumer2014r], building in pre-written blocks of executable code that demonstrate key programming concepts for students and including incomplete or blank code blocks that students can use to write their own code.
